---
title: "Who let the Dogs out?"
author: "Jin Chen"
date: "7/30/2020"
output: github_document
---

#Purpose
To answer the question "Which county has the most dangerous dog incidents in PA?" by analyzing a dataset from the Pennsylvania government on dangerous dog incidents from 1996 to 2019.

#Content
I will use the RSocrata package to import the data.

```{r Load Data}
library(tidyverse)
library(ggplot2)
library(RSocrata)
library(ggthemes)
library(ezknitr)

#set the public URL for the data set 
url <- "https://data.pa.gov/Public-Safety/Dangerous-Dogs-1996-Current-County-Agriculture/3fcn-e5dk"

#call the API using RSocrata
dan_dogs_raw <- read.socrata(url)

#show the first few obs
head(dan_dogs_raw)

```

#Structure
I create a separate df from the raw data. I rename a column to make it more accessible. Then I find the count of incidents that happened in each county. 

```{r Table Structure}
#by county
dan_dogs_county <- dan_dogs_raw %>% 
  rename(county = county_description) %>% 
  mutate( year = lubridate::year(file_date)) %>% #create a year column
  filter(year != 2020) %>% #filter out entries with year of 2020
  group_by(county) %>% 
  summarize(count = n()) %>% #count the number of incidents in each county
  select(county, count) %>% 
  mutate(county = fct_reorder(county, count)) %>%  #order county by count for graphing purposes
  arrange(desc(count)) %>% 
  top_n(10)

#Create a column to distinguish the #1 county of incidents with the rest. (could use gghighlight as an alternative)
dan_dogs_county <- dan_dogs_county %>% 
  mutate(county_highlight =
           ifelse(county == "ALLEGHENY",
                  "Allegheny",
                  "Other Counties"))


```
#Formatting
```{r Dangerous Dog Incidents Graph}
#by county
p <- ggplot(data = dan_dogs_county,
            mapping = aes(x = county,
                          y = count))
p + geom_col(aes(fill = county_highlight)) +
    coord_flip() +
    theme_tufte() +
    theme(legend.position = "none") +
    scale_fill_manual(values = c("dark blue", "light grey")) +
    labs(title = "Who Let the Dogs Out?",
       caption = "Data from 1996 - 2019",
       y = "Dangerous Dog Incidents",
       x = "") +
    theme(plot.title = element_text(hjust = 0.5)) +
    geom_text(aes(label = count), vjust = 0, hjust = -0.5, size = 3)
```

```{r Knit Output}
ezknit(file = "R/dan_dogs.Rmd", out_dir = "outputs/", keep_html = FALSE, fig_dir = "figures")

```

#Conclusion

Thus conclude today's daily data visualization! We have found that the PA county with the most dangerous dog incidents is Allegheny followed by York and Westmoreland with Bucks closely behind. 


